A Plasma Torus Around a Young Low-Mass Star

### 2025-05-28 paper/ms.tex
update

\begin{abstract}
  A small fraction of red dwarfs younger than $\sim$100 million years
  show structured, periodic optical light curves suggestive of
  transiting opaque material that corotates with the star.  However, the
  composition, origin, and even the existence of this material are
  uncertain. The main alternative hypothesis is that these complex
  periodic variables (CPVs) are explained by complex distributions of
  bright or dark regions on the stellar surfaces.  Here, we present
  time-series spectroscopy and photometry of a rapidly-rotating
  ($P$=3.9\,hr) CPV, TIC\,141146667.  The spectra show sinusoidal
  time-varying H$\alpha$ emission at twice to four times the star's
  equatorial velocity, providing direct evidence for cool
  ($\lesssim$10$^4$\,K) plasma clumps trapped in corotation around a
  CPV.  These data support the idea that young, rapidly-rotating M
  dwarfs can sustain warped \added{tori of cool plasma, similar to
  other rapidly-rotating magnetic stars.}
  \added{Outstanding questions include whether dust clumps 
  in these plasma tori explain CPV light curves, and
  whether the tori originate from the star or are fed by external
  sources.}
  %It remains unclear whether these structures originate intrinsically
  %from the star or are fed by external sources. Likewise,
  %the mechanism by which their optically thick clumps arise is still
  %unknown.
  Rough estimates suggest $\gtrsim$10\% of M dwarfs host
  similar structures during their early lives.
\end{abstract}

@@

\caption{Selected system parameters for TIC\,141146667.}

@@

In a follow-up study, \citet{Townsend2008} synthesized light curves for
these systems, assuming that the optical depth scaled linearly with
plasma density.  They found that W-shaped eclipses, similar to those seen for
TIC\,141146667, can occur when the spin and magnetic axes are moderately
(15-45$^\circ$) misaligned.  The movies associated with their work are
available
online\footnote{\url{http://user.astro.wisc.edu/\~townsend/static.php?ref=rrm-movies\#Download\_Bundles}
last accessed 23 April 2025.}; for a $\omega/\omega_{\rm
c}$$\approx$0.5, edge-on system like TIC\,141146667, the photometric
eclipses have the correct shape, and the H$\alpha$ emission similarly
exhibits double-peaked behavior.  The plasma in this model accumulates
at antipodes 180$^\circ$ apart because the deepest local minima in the
effective potential exist along the line of intersection between the
rotational and magnetic planes (see Equation~22 of
\citealt{Townsend2005}).

To summarize, the evidence for warped plasma tori
in TIC\,141146667 and CPVs more broadly is that 
Figure~\ref{fig:spec} shows two plasma clumps separated
by 180$^\circ$ in phase; the warped-torus model predicts both
``W'' shaped photometric eclipses and double-peaked H$\alpha$
morphology at quadrature; and CPV dips last 5-50\% of each cycle,
indicating an azimuthally extended distribution of material.
\added{However, two challenges remain. First, the model predicts {\it
two} W-shaped eclipses per cycle.  Second, it predicts that the
antipodal clumps should lie at equal distances from the star.
Different clump opacities or non-dipolar magnetic fields might resolve
these discrepancies.  Overall, the evidence suggests a warped torus is
a good first approximation---one that will need refinement in future
work.}

@@

It is currently unclear what, if any, relationship CPVs have to the
close-in rocky exoplanets that exist around most M dwarfs
\citep{Dressing2015}.  However, a few percent of young M dwarfs show
the CPV phenomenon \citep{Rebull2020}, and our data show that some of
the complex photometric features occur when clumps of circumstellar
material transit the star.  The geometric correction implies
that an appreciable minority ($\gtrsim$10\%) of young M dwarfs---the
rapidly rotating ones with centrifugal magnetospheres---host
circumstellar environments similar to the CPVs.  Future studies that
combine spectroscopic, polarimetric, and multi-wavelength
observations, along with magnetohydrodynamic modeling, will be key to
understanding the complex environments of these young
stars.

@@

{\it Radial Velocity}---We measured the radial velocities of
TIC\,141146667 from our HIRES spectra using a \added{custom }pipeline.
Our method is based on template-matching against synthetic
\added{PHOENIX }spectra \citep{Husser2013}, \added{and is }calibrated
\added{against }the standard stars described by \citet{Chubak2012}.
We used velocity standards spanning G2-M4, irrespective of rotation
rate.  We included velocity corrections due to Earth's motion around
the solar system barycenter and due to Earth's daily rotation about
its axis, calculated using \texttt{barycorrpy} \citep{Kanodia2018}.
Our analysis code reproduces the systemic velocities of known velocity
standards \citep{Chubak2012} with an RMS of 0.66\,\kms.

For TIC\,141146667, we measured the radial velocities using regions
near the K~I (7700\,\AA) resonance line and three TiO bandheads
(5160\,\AA, 5450\,\AA, and 5600\,\AA).  We selected these regions
because they provided the best matches between the synthetic and
observed spectra.  We then averaged the resulting redshift
measurements over each order to produce the final measurement.  We
used the scatter of resulting velocity measurements between orders to
assign the RV uncertainty at each epoch.  The uncertainty-weighted
mean systemic velocity over all epochs on 17~February~2024 was
$\gamma$=0.6$\pm$1.5\,\kms.  The relative radial velocities about this
mean are listed in Table~\ref{tab:rv}.

@@

{\it No Evidence For Binarity}---Any periodicity in the radial
velocity time-series is ruled out at the rotation period for
semi-amplitudes above 2.85\,\kms\ (at 3$\sigma$ confidence).  This
sets an upper limit on the mass of any putative companion at the four
hour period of $m \sin i $$<$2.4\,$M_{\rm Jup}$.  Regarding possible
companions at wider separations, the Gaia DR3 renormalized unit weight
error (RUWE), a proxy for the goodness of fit for a single-source
astrometric model to the Gaia astrometry, is 1.23, within the usual
range for apparently single sources.  There are no resolved companions
in the Gaia DR3 point source catalog.  Finally, we checked the TESS
light curve for evidence of secondary photometric periods by
subtracting the mean CPV signal over each sector and performing a
phase-dispersion minimization analysis
\citep{Stellingwerf1978,2021zndo...1011188B}.  There were no secondary
periods in the TESS data.  Previous work \citep{Bouma2024} has shown
that $\approx$30\% of CPVs show evidence for excess noise above the
Gaia single-source astrometric model, and $\approx$40\% of CPVs show
evidence for unresolved binary companions based on the presence of
secondary photometric periods.  This agrees with analyses showing that
multi-periodic low-mass stars are generally unresolved binaries
\citep{Tokovinin2018}.  Overall, the CPV binary fraction seems
consistent with that for field M dwarfs \citep{Winters2019}, pointing
to a weak or non-existent connection between the CPV phenomenon and
(wide) binarity.  For TIC\,141146667 specifically, we find no
evidence for stellar multiplicity. %, our data are minimally constraining
%for the scenario of a low-luminosity companion
%($F_1$/$F_2$$\lesssim$0.1) with apparent separation below 0$\farcs$1.

@@

To implement this model, we assumed that the ``inner'' ($K_{\rm
inner}$$\approx$2.5\,$v_{\rm eq}$) clump would be well-fit by a sum of
two gaussians because it is visually double-peaked in the raw data
from $\phi$=0.15-0.35 and $\phi$=0.65-0.85 (Figure~\ref{fig:spec}b).
We assumed that the ``outer'' ($K_{\rm outer}$$\approx$3.9\,$v_{\rm
eq}$) clump would be better fit by a single gaussian, based on its
behavior from $\phi$=0.6-0.9.  Each gaussian component has three free
parameters at each spectral epoch: the mean $\mu$, standard deviation
$\sigma$, and amplitude $A$.  We labeled the inner component's two
gaussians $i=\{ 0, 1 \}$, and the single outer component as $i=2$.
Given the complexity of the line profile data
(Figure~\ref{fig:halphamodel} left column), the likelihood function
for this model is multimodal.  We therefore imposed the prior
constraints that $A_i \sim \mathcal{U}[0,1]$,
$\sigma_i/v_{\rm eq} \sim \mathcal{U}[0,1]$, and further assumed
$\mu_i(t) \sim
\mathcal{U}[
  K_{\rm inner}\sin(\phi(t)) - v_{\rm eq},
  K_{\rm inner}\sin(\phi(t)) + v_{\rm eq}
]$
for the inner two components, and
$\mu_2(t) \sim
\mathcal{U}[
  K_{\rm outer}\sin(\phi(t) + \pi) - 2v_{\rm eq},
  K_{\rm outer}\sin(\phi(t) + \pi) + 2v_{\rm eq}
]$
for the outer component.  This prior on the means mitigates
multimodality in the likelihood by requiring the mean velocity of each
component to be within a one or two $v_{\rm eq}$ of the time-variable
sinusoid suggested by visual inspection.  We fitted each component to
the data {\it independently} using scipy's non-linear least squares
\texttt{curve\_fit} implementation \citep{Virtanen2020}, and scaled the
resulting parameter covariance matrix by a constant factor to match
the sample variance of the residuals.
\added{The results are shown in the middle columns of
Figure~\ref{fig:halphamodel}, and a table of the fitted means,
amplitudes, and standard deviations for each component is available
upon request.}

@@

Accounting for these caveats, we used the \added{best-fit parameters
from the multi-gaussian model }to quantify the orbital periods and
velocity semi-amplitudes of the clumps.  Figure~\ref{fig:orbits}a
shows the results assuming circular orbits; considering the Bayesian
information criterion, we found no reason to prefer eccentric orbits.
By visually inspecting Figure~\ref{fig:halphamodel}, we
excluded epochs where our gaussian profile fitting failed to either
detect or else adequately represent the circumstellar emission.  We
then used NumPyro to sample the Gaussian likelihood for a circular
orbit with the NUTS algorithm \citep{Phan2019}.  We used the
measurement uncertainties from each estimated mean radial velocity
value and included an additional jitter term in quadrature.  This
procedure yielded orbital periods and semi-amplitudes of
$P_0$=$3.92\pm0.03$\,hr, $K_0/v_{\rm eq}$=2.07$\pm$0.04;
$P_1$=$3.92\pm0.06$\,hr, $K_1/v_{\rm eq}$=2.88$\pm$0.10;
and $P_2$=$3.88\pm0.20$\,hr, $K_1/v_{\rm eq}$=3.88$\pm$0.25.
The periods for the inner double-peaked clump are therefore consistent
with the photometric $3.930\pm0.001$\,hr period within a precision of
two minutes.  The ``period'' for the outer clump is ambiguous because
the H$\alpha$ data only support the idea of a periodic orbit of
material well-fit by gaussian emission from $\phi$$\approx$0.5-1.0.
From $\phi$=0-0.2, there is no detectable emission, and from
$\phi$=0.2-0.4, the emission spans 1-4\,$v_{\rm eq}$ without taking a
clear gaussian shape.  While the outer-most edge of this
$\phi$=0.2-0.4 emission provides a plausible match to the expectation
of a circular orbit, the idea of invoking a particular functional form
for this component seems fine-tuned.  We instead emphasize that
although this emission is present, its variability in time is
inconsistent with the idea of a stable clump of material.  Additional
observations would be needed to conclusively determine whether or not
this component of the system is long-lived.

@@

{\it Physical dimensions of the emitting region}---The measured
velocity widths from the circumstellar emission contain information
about the size of the emitting region via the condition for rigid
corotation.  Consider a clump in cylindrical coordinates with
arbitrary radial extent $r$, azimuthal extent $\ell$, and height $z$.
A range of shapes, including an ``arc'' with $\ell \gg r$, a ``spoke''
with $r \gg \ell$, and a ``blob'' with $r \approx \ell \approx z$ are
all a priori possible.  However, at quadrature, the observed velocity
width of emission is sensitive to the radial extent of the
circumstellar material.  At mid-transit, the observed velocity width
is sensitive to the azimuthal extent.  An arc configuration would
minimize the observed velocity width $\sigma$ at quadrature, and
maximize it during transit, with $\gtrsim$100\,\kms\ variations in
between.  This is not observed.   The arc geometry, and by a similar
argument the spoke geometry, can thus be discarded.

At quadrature, the inner clumps show $\sigma_i$$\approx$0.24\,$v_{\rm
eq}$, implying that 68\% of the emission comes from a volume with
length in the radial dimension $r_i$=$2\sigma_i /
\Omega$$=$0.48\,$R_\star$, and that 95\% of the emission comes from
within 0.96\,$R_\star$.  These values have relative uncertainties of
$\approx$5\%, based on the uncertainties in the measured velocity
dispersions.  The two inner clumps are centered at orbital distances
of 2.07\,$R_\star$ and 2.88\,$R_\star$.  There is therefore physical
overlap in their spatial distributions.  These two clumps could in
fact be a single clump with an optically thick H$\alpha$ line core.
Regardless of this nuance, the implication is that the full length in
the radial dimension of these two inner emitting clumps is
approximately equal to the star's diameter.  At mid-transit, these
inner clumps have a similar velocity dispersion, although with
greater uncertainty due to the differences between the $\phi$=0 and
$\phi$=1 transits (see Figure~\ref{fig:halphamodel}).
This suggests a 1$\sigma$ emission contour in the azimuthal dimension
with a length of $\ell$$\approx$0.5\,$R_\star$.

The vertical height of the emitting region is less constrained because
the system is consistent with being viewed edge-on.  However, one
constraint does come if one assumes that the TESS transit depth scales
with the projected H$\alpha$ emission area $\ell z$.  
More specifically, one can evaluate an ``effective area'' blocked by a
two-dimensional gaussian blob passing over a star by integrating the
local gaussian weight over the stellar surface.  For instance, if
$\ell$$\approx$$z$$\approx$0.24\,$R_\star$, then
$\iint \exp \left(
  -\left( x^2/2\sigma_x^2 + y^2/2\sigma_y^2 \right)
\right) {\rm d}x \, {\rm d}y$
suggests 11.5\% of the star being ``blocked'', a geometric factor
which would need to be in turn multiplied by an unknown opacity factor
to produce the observed transit depth ($\delta$$\approx$5\%).  In
general though, there is a degeneracy between $z$ and this opacity
factor; larger vertical heights are allowed for lower optical depths
in absorption, and vice-versa.  This constraint also implicitly
assumes that the optically thick material is well-mixed with the
hydrogen, which may not be accurate.

@@

The physical conditions inside the plasma clumps, in particular the
hydrogen number density, plasma temperature, ionization fraction, and
magnetic field strength can be estimated from the available data.  We
caution that the following estimates are order of magnitude
calculations that assume a simple uniform-density plasma: detailed
considerations of radiative transfer are a worthy topic for future
work, but are beyond the scope of this article.

Circumstellar H$\alpha$ emission might be sourced either from resonant
scattering of stellar H$\alpha$ photons, or from radiative
recombination.  We neglect scattering because Figure~\ref{fig:spec}
shows the amplitude of the
circumstellar H$\alpha$ emission varying by a factor of $\approx$5, in
a manner uncorrelated with any variability in the chromospheric line
core.  The volume emissivity under case B recombination can be written
\begin{equation} j_{\rm H\alpha} = n_{\rm e} n_{\rm p} \alpha^{\rm
eff}_{\rm H\alpha} h \nu_{\rm H\alpha}, \end{equation} where $n_{\rm
e}$ and $n_{\rm p}$ are the electron and proton densities, and
$\alpha^{\rm eff}_{\rm H\alpha}$ is the effective recombination
coefficient, defined to include all recombination routes that produce
an H$\alpha$ photon.  For hydrogen with temperatures between
1,000-10,000\,K, $\alpha^{\rm eff}_{\rm H\alpha}$ is typically on the
order of $10^{-12}$ to $10^{-13}$\,cm$^3$\,s$^{-1}$
\citep{Hummer1987,Draine2011}.  Neglecting the effects of atoms other
than hydrogen, we can assume an ionization fraction $x$, such that
$n_{\rm e} = n_{\rm p} = x n_{\rm H}$, for $n_{\rm H}$ the hydrogen
number density.  Let $L_{\rm H\alpha} = j_{\rm H\alpha} V$, for $V$
the volume of the emitting hydrogen.  The luminosity of circumstellar
hydrogen emission, $L_{\rm H\alpha}$, is an observable: our SED
fitting routine yields $L_\star$$\approx$0.012\,$L_\odot$, which
implies that the stellar H$\alpha$ line radiates at
$\approx$1.0$\times$$10^{28}$erg\,s$^{-1}$.  The luminosity of the
clumps $L_{\rm H\alpha}$ are of order one tenth that of the star.  If
we approximate the emitting volume as a homogeneous sphere of radius
$r$, we can write
\begin{align}
  n_{\rm H} &= 
  1 \cdot 10^{11}\,{\rm cm}^{-3}
  \left(
    \frac{0.5}{x}
  \right)
  \left( 
    \frac{ L_{\rm H\alpha} }{ 10^{27}\,{\rm erg\,s^{-1}} }
    \frac{ 10^{-13}\,{\rm cm^{3}\,s^{-1}} }{ \alpha^{\rm eff}_{\rm H\alpha} }
  \right)^{1/2}
  \left(
    \frac{ 0.1\,R_\odot }{ r }
  \right)^{3/2}.
  \label{eq:numdensity}
\end{align}
For a uniform density clump, this suggests a total gas mass of $M_{\rm
gas}\approx 2\times10^{17}$\,g\added{, which is similar to masses derived for
plasma clumps around other cool stars \citep{VillarrealDAngelo2019}.}.  We emphasize that
Equation~\ref{eq:numdensity} in intended to provide only an order of
magnitude estimate for the number density implied by the observed
H$\alpha$ emission.  In detail, the effective recombination rate and
the ionization fraction each vary with density and temperature; a more
thorough estimate would iteratively solve the equations of detailed
balance and radiative transfer (e.g.~Figure~8 of \citealt{CollierCameron1989}),
and potentially also consider departures from local thermodynamic equilibrium.

>>>>

\begin{abstract}
  A small fraction of red dwarfs younger than $\sim$100 million years
  show structured, periodic optical light curves suggestive of
  transiting opaque material that corotates with the star.  However, the
  composition, origin, and even the existence of this material are
  uncertain. The main alternative hypothesis is that these complex
  periodic variables (CPVs) are explained by complex distributions of
  bright or dark regions on the stellar surfaces.  Here, we present
  time-series spectroscopy and photometry of a rapidly-rotating
  ($P$=3.9\,hr) CPV, TIC\,141146667.  The spectra show sinusoidal
  time-varying H$\alpha$ emission at twice to four times the star's
  equatorial velocity, providing direct evidence for cool
  ($\lesssim$10$^4$\,K) plasma clumps trapped in corotation around a
  CPV.  These data support the idea that young, rapidly-rotating M
  dwarfs can sustain warped \added{tori of cool plasma, similar to
  other rapidly-rotating magnetic stars.}
  \added{Outstanding questions include whether dust clumps 
  in these plasma tori explain CPV light curves, and
  whether the tori originate from the star or are fed by external
  sources.}
  Rough estimates suggest $\gtrsim$10\% of M dwarfs host
  similar structures during their early lives.
\end{abstract}

@@

  \caption{Selected properties of TIC\,141146667.}

@@

In a follow-up study, \citet{Townsend2008} made synthetic light
curves, assuming that the optical depth scaled linearly with plasma
density.  They found that W-shaped eclipses, similar to those seen for
TIC\,141146667, can occur when the spin and magnetic axes are
moderately (15-45$^\circ$) misaligned.  The movies associated with
their work are available
online\footnote{\url{http://user.astro.wisc.edu/\~townsend/static.php?ref=rrm-movies\#Download\_Bundles}
last accessed 23 April 2025.}; for a $\omega/\omega_{\rm
c}$$\approx$0.5, edge-on system like TIC\,141146667, the photometric
eclipses have the correct shape, and the H$\alpha$ emission similarly
exhibits double-peaked behavior.  The plasma in this model accumulates
at antipodes 180$^\circ$ apart because the deepest local minima in the
effective potential exist along the line of intersection between the
rotational and magnetic planes (see Equation~22 of
\citealt{Townsend2005}).

To summarize, the evidence for warped plasma tori
in TIC\,141146667 and CPVs more broadly is that 
Figure~\ref{fig:spec} shows two plasma clumps separated
by 180$^\circ$ in phase; the warped-torus model predicts both
``W'' shaped photometric eclipses and double-peaked H$\alpha$
morphology at quadrature; and CPV dips last 5-50\% of each cycle,
indicating an azimuthally extended distribution of material.
\added{However, two challenges remain. First, the model predicts {\it
two} W-shaped eclipses per cycle.  Second, it predicts that the
antipodal clumps should lie at equal distances from the star.
Different clump opacities or non-dipolar magnetic fields might resolve
these discrepancies.  Overall, the evidence suggests a warped torus is
a good first approximation, but one that will need refinement in future
work.}

@@

It is currently unclear what, if any, relationship CPVs have to the
close-in rocky exoplanets that exist around most M dwarfs
\citep{Dressing2015}.  However, a few percent of young M dwarfs show
the CPV phenomenon \citep{Rebull2020}, and our data show that some of
the flux dips occur when clumps of circumstellar
material transit the star.  The geometric correction implies
that an appreciable minority ($\gtrsim$10\%) of young M dwarfs---the
rapidly rotating ones with centrifugal magnetospheres---host
circumstellar environments similar to the CPVs.  Future studies that
combine spectroscopic, polarimetric, and multi-wavelength
observations, along with magnetohydrodynamic modeling, will be key to
understanding the complex environments of these young
stars.

@@

{\it Radial Velocity}---We measured the radial velocities of
TIC\,141146667 \added{(Gaia DR3 860453786736413568) }from our HIRES
spectra using a \added{custom }pipeline.  Our method was based on
template-matching against synthetic \added{PHOENIX }spectra
\citep{Husser2013}, \added{and was }calibrated \added{against }the
standard stars described by \citet{Chubak2012}.  We used velocity
standards spanning G2-M4, irrespective of rotation rate.  We included
velocity corrections due to Earth's motion around the solar system
barycenter and due to Earth's daily rotation about its axis,
calculated using \texttt{barycorrpy} \citep{Kanodia2018}.  Our
analysis code reproduced the systemic velocities of known velocity
standards \citep{Chubak2012} with an RMS of 0.66\,\kms.

For TIC\,141146667, we measured the radial velocities using regions
near the K~I (7700\,\AA) resonance line and three TiO bandheads
(5160\,\AA, 5450\,\AA, and 5600\,\AA).  We selected these regions
because they provided the best matches between the synthetic and
observed spectra.  We averaged the redshift
measurements over each order, and
used the scatter of resulting velocity measurements between orders to
assign the RV uncertainty at each epoch.  The uncertainty-weighted
mean systemic velocity over all epochs on 17~February~2024 was
$\gamma$=0.6$\pm$1.5\,\kms.  The relative radial velocities about this
mean are listed in Table~\ref{tab:rv}.

@@

{\it No Evidence For Binarity}---Any periodicity in the radial
velocity time-series is ruled out at the rotation period for
semi-amplitudes above 2.85\,\kms\ (at 3$\sigma$ confidence).  This
sets an upper limit on the mass of any putative companion on a four
hour orbit of $m \sin i $$<$2.4\,$M_{\rm Jup}$.  Regarding possible
companions at wider separations, the Gaia DR3 renormalized unit weight
error (RUWE) is 1.23, within the usual
range for apparently single sources.  There are no resolved companions
in the Gaia DR3 point source catalog.  Finally, we checked the TESS
light curve for evidence of secondary photometric periods by
subtracting the mean CPV signal over each sector and performing a
phase-dispersion minimization analysis
\citep{Stellingwerf1978,2021zndo...1011188B}.  There were no secondary
periods in the TESS data.  Previous work \citep{Bouma2024} has shown
that $\approx$30\% of CPVs show evidence for binarity from RUWE, and $\approx$40\% of CPVs show
evidence for binarity based on the presence of
secondary photometric periods.  This agrees with analyses showing that
multi-periodic low-mass stars are usually unresolved binaries
\citep{Tokovinin2018}.  Overall, the CPV binary fraction seems
consistent with that of field M dwarfs \citep{Winters2019}, pointing
to a weak or non-existent connection between the CPV phenomenon and
(wide) binarity.  For TIC\,141146667 specifically, we find no
evidence for stellar multiplicity. %, our data are minimally constraining
%for the scenario of a low-luminosity companion
%($F_1$/$F_2$$\lesssim$0.1) with apparent separation below 0$\farcs$1.

@@

To implement this model, we assumed that the ``inner'' ($K_{\rm
inner}$$\approx$2.5\,$v_{\rm eq}$) clump would be well-fit by a sum of
two gaussians because it is visually double-peaked in the raw data
from $\phi$=0.15-0.35 and $\phi$=0.65-0.85 (Figure~\ref{fig:spec}b).
We assumed that the ``outer'' ($K_{\rm outer}$$\approx$3.9\,$v_{\rm
eq}$) clump would be better fit by a single gaussian, based on its
behavior from $\phi$=0.6-0.9.  Each gaussian component has three free
parameters at each spectral epoch: the mean $\mu$, standard deviation
$\sigma$, and amplitude $A$.  We labeled the inner component's two
gaussians $i=\{ 0, 1 \}$, and the single outer component as $i=2$.
Given the complexity of the line profile data
(Figure~\ref{fig:halphamodel} left column), the likelihood function
for this model is multimodal.  We therefore imposed the prior
constraints that $A_i \sim \mathcal{U}[0,1]$,
$\sigma_i/v_{\rm eq} \sim \mathcal{U}[0,1]$, and further assumed
$\mu_i(t) \sim
\mathcal{U}[
  K_{\rm inner}\sin(\phi(t)) - v_{\rm eq},
  K_{\rm inner}\sin(\phi(t)) + v_{\rm eq}
]$
for the inner two components, and
$\mu_2(t) \sim
\mathcal{U}[
  K_{\rm outer}\sin(\phi(t) + \pi) - 2v_{\rm eq},
  K_{\rm outer}\sin(\phi(t) + \pi) + 2v_{\rm eq}
]$
for the outer component.  This prior on the means mitigates
multimodality in the likelihood by requiring the mean velocity of each
component to be within a one or two $v_{\rm eq}$ of the time-variable
sinusoid suggested by visual inspection.  We fitted each component to
the data {\it independently} using scipy's non-linear least squares
\texttt{curve\_fit} implementation \citep{Virtanen2020}, and scaled the
resulting parameter covariance matrix by a constant factor to match
the sample variance of the residuals.
\added{The middle columns of
Figure~\ref{fig:halphamodel} show the results; a table of the fitted means,
amplitudes, and standard deviations for each component is available
upon request.}

@@

Accounting for these caveats, we used the \added{best-fit parameters
from the multi-gaussian model }to quantify the orbital periods and
velocity semi-amplitudes of the clumps.  Figure~\ref{fig:orbits}a
shows the results assuming circular orbits; considering the Bayesian
information criterion, we found no reason to prefer eccentric orbits.
By visually inspecting Figure~\ref{fig:halphamodel}, we
excluded epochs where our gaussian profile fitting failed to either
detect or else adequately represent the circumstellar emission.  We
then used NumPyro to sample the Gaussian likelihood for a circular
orbit with the NUTS algorithm \citep{Phan2019}.  We used the
measurement uncertainties from each estimated mean radial velocity
value and included an additional jitter term in quadrature.  This
procedure yielded orbital periods and semi-amplitudes of
$P_0$=$3.92\pm0.03$\,hr, $K_0/v_{\rm eq}$=2.07$\pm$0.04;
$P_1$=$3.92\pm0.06$\,hr, $K_1/v_{\rm eq}$=2.88$\pm$0.10;
and $P_2$=$3.88\pm0.20$\,hr, $K_2/v_{\rm eq}$=3.88$\pm$0.25.
The periods for the inner double-peaked clump are therefore consistent
with the photometric $3.930\pm0.001$\,hr period within a precision of
two minutes.  The ``period'' for the outer clump is ambiguous because
the H$\alpha$ data only support the idea of a periodic orbit of
material well-fit by gaussian emission from $\phi$$\approx$0.5-1.0.
From $\phi$=0-0.2, there is no detectable emission, and from
$\phi$=0.2-0.4, the emission spans 1-4\,$v_{\rm eq}$ without taking a
clear gaussian shape.  While the outer-most edge of this
$\phi$=0.2-0.4 emission provides a plausible match to the expectation
of a circular orbit, the idea of invoking a particular functional form
for this component seems fine-tuned.  We instead emphasize that
although this emission is present, its variability in time is
inconsistent with the idea of a stable clump of material.  Additional
observations would be needed to conclusively determine whether or not
this component of the system is long-lived.

@@

{\it Physical dimensions of emitting region}---The H$\alpha$
velocity widths in Figure~\ref{fig:halphamodel} constrain
the size of the emitting region via the condition for rigid
corotation.  Consider a clump in cylindrical coordinates with
arbitrary radial extent $r$, azimuthal extent $\ell$, and height $z$.
A range of shapes, including an ``arc'' with $\ell \gg r$, a ``spoke''
with $r \gg \ell$, and a ``blob'' with $r \approx \ell \approx z$ are
all a priori possible.  However, at quadrature, the observed velocity
width of emission is sensitive to the radial extent of the
circumstellar material.  At mid-transit, the observed velocity width
is sensitive to the azimuthal extent.  An arc configuration would
minimize the observed velocity width $\sigma$ at quadrature, and
maximize it during transit, with $\gtrsim$100\,\kms\ variations in
between.  Since this is not observed, the arc geometry, and by a similar
argument the spoke geometry, can be discarded.

At quadrature, the inner clumps show $\sigma_i$$\approx$0.24\,$v_{\rm
eq}$, implying that 68\% of the emission comes from a volume with
length in the radial dimension $r_i$=$2\sigma_i /
\Omega$$=$0.48\,$R_\star$, and that 95\% of the emission comes from
within 0.96\,$R_\star$.  These values have relative uncertainties of
$\approx$5\%, based on the uncertainties in the measured velocity
dispersions.  The two inner clumps are centered at orbital distances
of 2.07\,$R_\star$ and 2.88\,$R_\star$.  There is therefore physical
overlap in their spatial distributions.  These two clumps could in
fact be a single clump with an optically thick H$\alpha$ line core.
Regardless, the implication is that the full length in the radial
dimension of these two inner emitting clumps is approximately equal to
the star's diameter.  At mid-transit, these
inner clumps have a similar velocity dispersion, although with
greater uncertainty due to the differences between the $\phi$=0 and
$\phi$=1 transits (see Figure~\ref{fig:halphamodel}).
This suggests a 1$\sigma$ emission contour in the azimuthal dimension
with a length of $\ell$$\approx$0.5\,$R_\star$.

The vertical height of the emitting region is less constrained because
the system is consistent with being viewed edge-on.  However, one
possible constraint follows by assuming that the TESS transit depth scales
with the projected H$\alpha$ emission area $\ell z$.  
More specifically, one can evaluate an ``effective area'' blocked by a
two-dimensional gaussian blob passing over a star by integrating the
local gaussian weight over the stellar surface.  For instance, if
$\ell$$\approx$$z$$\approx$0.24\,$R_\star$, then
$\iint \exp \left(
  -\left( x^2/2\sigma_x^2 + y^2/2\sigma_y^2 \right)
\right) {\rm d}x \, {\rm d}y$
suggests 11.5\% of the star being ``blocked'', a geometric factor
which would need to be in turn multiplied by an unknown opacity factor
to produce the observed transit depth ($\delta$$\approx$5\%).  
There is a degeneracy between $z$ and this opacity
factor; larger vertical heights are allowed for lower optical depths
in absorption, and vice-versa.  This constraint also implicitly
assumes that the optically thick material is well-mixed with the
hydrogen, which may not be accurate.

@@

The hydrogen number density, plasma temperature, and magnetic field
strength inside the clumps can be estimated from the available data.  
The following order-of-magnitude calculations assume a simple
uniform-density plasma in a spherical geometry: more careful
considerations of radiative transfer are a worthy topic for future
work.

Circumstellar H$\alpha$ emission could come either from resonant
scattering of stellar H$\alpha$ photons, or from radiative
recombination.  We neglect scattering because Figure~\ref{fig:spec}
shows the
circumstellar H$\alpha$ emission varying by a factor of $\approx$5
while the chromospheric line core is stable.  The volume emissivity
under case B recombination can be written \begin{equation}
  j_{\rm H\alpha} = n_{\rm e} n_{\rm p} \alpha^{\rm eff}_{\rm H\alpha} h \nu_{\rm H\alpha},
\end{equation} where $n_{\rm e}$ and $n_{\rm p}$ are the electron and
proton densities, and $\alpha^{\rm eff}_{\rm H\alpha}$ is the
effective recombination coefficient, defined to include all
recombination routes that produce an H$\alpha$ photon.  For hydrogen
with temperatures between 1,000-10,000\,K, $\alpha^{\rm eff}_{\rm
H\alpha}$ is typically $10^{-12}$ to
$10^{-13}$\,cm$^3$\,s$^{-1}$ \citep{Hummer1987,Draine2011}.
Neglecting atoms other than hydrogen, we can assume an
ionization fraction $x$, such that $n_{\rm e} = n_{\rm p} = x n_{\rm
H}$, for $n_{\rm H}$ the hydrogen number density.  Let $L_{\rm
H\alpha} = j_{\rm H\alpha} V$, for $V$ the volume of the emitting
hydrogen.  The luminosity of circumstellar hydrogen emission, $L_{\rm
H\alpha}$, is an observable: our SED fitting routine yields
$L_\star$$\approx$0.012\,$L_\odot$, which implies that the stellar
H$\alpha$ line radiates at $\approx$1.0$\times$$10^{28}$erg\,s$^{-1}$.
The luminosity of the clumps $L_{\rm H\alpha}$ are of order one tenth
that of the star.  If we approximate the emitting volume as a
homogeneous sphere of radius $r$, we can write
\begin{align}
  n_{\rm H} &= 
  1 \cdot 10^{11}\,{\rm cm}^{-3}
  \left(
    \frac{0.5}{x}
  \right)
  \left( 
    \frac{ L_{\rm H\alpha} }{ 10^{27}\,{\rm erg\,s^{-1}} }
    \frac{ 10^{-13}\,{\rm cm^{3}\,s^{-1}} }{ \alpha^{\rm eff}_{\rm H\alpha} }
  \right)^{1/2}
  \left(
    \frac{ 0.1\,R_\odot }{ r }
  \right)^{3/2}.
  \label{eq:numdensity}
\end{align}
For a uniform density clump, this suggests a total gas mass of $M_{\rm
gas}\approx 2\times10^{17}$\,g\added{, which is similar to masses derived for
plasma clumps around other cool stars \citep{VillarrealDAngelo2019}}.  We emphasize that
Equation~\ref{eq:numdensity} in intended to provide only an order of
magnitude estimate for the number density implied by the observed
H$\alpha$ emission.  In detail, the effective recombination rate and
the ionization fraction each vary with density and temperature; a more
thorough estimate would iteratively solve the equations of detailed
balance and radiative transfer (e.g.~Figure~8 of \citealt{CollierCameron1989}),
and potentially also consider departures from local thermodynamic equilibrium.

Finally, a constraint on the magnetic field strength at the site of
the clump follows from the requirement that the magnetic pressure
exceed the thermal pressure, $B_{\rm c}^2 / 8\pi > n_{\rm H} k T$.
Although we do not know the exact plasma temperature, if it were
significantly beyond 1,000-10,000\,K, we would either fully ionize the
hydrogen, or not ionize enough of it.  The field strength at the clump
must therefore exceed
\begin{equation}
  B_{\rm c} \gtrsim 1\,{\rm G}
  \left(
  \frac{n_{\rm H}}{1\times10^{11}\,{\rm cm}^{-3}}
  \frac{T}{3000\,{\rm K}}
  \right)^{1/2}.
\end{equation}
Given that the average surface magnetic field strengths of low-mass
stars have been measured to span hundreds to thousands of Gauss
\citep{Donati2009,Kochukhov2021,Reiners2022}, this bound is easily met
at orbital distances of 2-4\,$R_\star$.
